---
output:
  word_document: default
---
--
title: "Statistical Foundations for Data Science - Assignment8 "
author: "Travis Deason"
date: "November 18th 2017"
output:  word_document
---

# MSDS 6371 UNIT 11 HW


```{r, echo=FALSE, message=FALSE}
rm( list = ls()); cat("\014")  # Clear environment
#install.packages('hmisc')
#install.packages('ggplot2')
#library(Hmisc)
library(ggplot2)
metdata <- read.csv('data/MetabolismDataProb26_2_2.csv')
names(metdata) <- sapply(names(metdata), tolower)
```
## Problem 26 Chapter 8 The Metabolic data has the average mass, metabolic rate and average lifespan of 95 different species of mammals.  Kleiber’s law states that the metabolic rate of an animal species, on average, is proportional to its mass raised to the power of ¾.  Judge the adequacy of this theory with these data.  
### Be sure and provide:
>* A Scatterplot with confidence intervals of the regression line and prediction intervals of the regression line.  In SAS and R!

```{r echo=TRUE, message=FALSE}
model <- lm(metab~mass, data=metdata)
ggplot(metdata$mass, model$residuals, )
```


>* A table showing the t-statistics and p-values for the significance of the regression parameters: β_(0 ) and β_1.   In SAS and R!
	The regression equation.
	Interpretation of the model given any transformation you may use.  Hint Hint ;) 
	A scatterplot of residuals. In SAS and R!
	A histogram of residuals with normal distribution superimposed.  (from SAS).
	Provide a measure of the amount of variation in the response that is accounted for by the explanatory variable.  Interpret this measure clearly indicating the units of the response and the explanatory variables.
	
	
	
	
	
	

## Problem 29 Chapter 8
The Autism data shows the prevalence of autism per 10,000 ten-year old children in the United States in each of five years.  Analyze the data to describe the change in the distribution of autism prevalence per year in this time period.  
	A Scatterplot with confidence intervals of the regression line and prediction intervals of the regression line.  In SAS OR R!
	A table showing the t-statistics and p-values for the significance of the regression parameters: β_(0 ) and β_1.  In SAS OR R!
	The regression equation.
	Interpretation of the model given any transformation you may use.  Hint Hint ;) 
	A scatterplot of residuals.  In SAS OR R!
	A histogram of residuals with normal distribution superimposed.  (from SAS).
	Provide a measure of the amount of variation in the response that is accounted for by the explanatory variable.  Interpret this measure clearly indicating the units of the response and the explanatory variables.

 
 
 
 
 


