---
output:
  word_document: default
---
--
title: "Statistical Foundations for Data Science - Assignment8 "
author: "Travis Deason"
date: "Oct 22nd 2017"
output:  word_document
---

# MSDS 6371 UNIT 10 HW
* These are the same data from last week’s HW.  Now, we are going to use them for simple linear regression.

```{r, echo=FALSE, message=FALSE}
rm( list = ls()); cat("\014")  # Clear environment
library(Hmisc)
library(ggplot2)
wheaters <- read.csv('data/MaleDisplayDataSet_2_2.csv')
names(wheaters) <- sapply(names(wheaters), tolower)
```
# UNIT 10: STAT 6371: Chapter 7
## 1. Black-eared wheatears are small birds of Spain and Morocco.  Males of the species demonstrate an exaggerated sexual display by carrying many heavy stones to nesting cavities.  This 35-gram bird transports, on average, 3.1 kg of stones per nesting season!  Different males carry somewhat different sized stones, prompting a study of whether larger stones may be a signal of higher health status.  M. Soler et al. calculated the average stone mass (g) carried by each of 21 male black-eared wheatears, along with T-cell response measurements reflecting their immune system strengths.  Analyze the data and write a statistical report (by answering the questions below); treat the T-cell as the response and the stone mass as the explanatory variable.

### a. You can find the data for this problem on 2DS. (Male Display Data Set.csv). Analyze the data providing at least the following: 

#### i. Provide a scatterplot with confidence intervals of the regression line and prediction intervals of the regression line.  Please do in SAS and R!

#### ii. Provide a table showing the t-statistics and p-values for the significance of the regression parameters: β_(0 ) and β_1. Please do in SAS and R!

#### iii. Using the output in (ii), show all 6 steps of each hypothesis test. 

#### iv. State the regression equation.

#### v. Interpret the slope and y-intercept in the model (regression equation). 

#### vi. Find and interpret the 95% confidence interval for the mean t-cell response conditional on a stone mass of 4.5 grams.  Please do in SAS and R!

#### vii. Find and interpret the 95% prediction interval for the predicted t-cell response given a stone mass of 4.5 grams.   Please do in SAS and R!
	
#### viii. Using the graphical method, find and interpret the calibration intervals for the t-cell response of 0.3. (Do this for both for mean t-cell response and for a single t-cell response. Please do in R as SAS does not provide calibration intervals! R: package investr) 

#### ix. Find the same calibration intervals analytically using the SE equations given in class and in the book (Version 3 page 194).

#### x. Provide a scatterplot of residuals. Please do in SAS and R!

#### xi. Provide a histogram of residuals with normal distribution superimposed (from SAS). 

```{r, echo=TRUE, message=FALSE}
#Perform Regression
model = lm(mass~tcell, data=wheaters)
#Extract Regression line coefficents
wheaters$resid <- model$residuals
slope <- model$coefficients['tcell']
intercept <- model$coefficients[1]
#Prediction intervals
pred.int =  predict(model, interval="prediction")
#Confidence intervals
conf.int =  predict(model, interval="confidence")
# Add predictions and CI to original data
wheaters$pred.lower <- pred.int[,2]
wheaters$pred.upper <- pred.int[,3]
wheaters$ci.upper <- conf.int[,2]
wheaters$ci.lower <- conf.int[,3]
## plot
ggplot(data=wheaters, aes(x=tcell, y=mass, main='Scatterplot of data with regression line and Confidence intervals')) + 
  geom_point(color= 'red') +
  geom_abline(intercept=intercept, slope=slope, color='black') +
  geom_ribbon(data=wheaters, aes(ymin= pred.lower, ymax= pred.upper), fill = "blue", alpha = 0.2) +
  geom_ribbon(data=wheaters, aes(ymin= ci.lower, ymax= ci.upper), fill = "cyan", alpha = 0.3) 
```



```{r, echo=TRUE, message=FALSE}
summary(model)
print(confint(model, level=.975))
# Find critical t-value for slope
qt(.975, 19)
qplot(model$residuals, geom="histogram", main='Histogram of Residuals')
plot(wheaters$tcell, wheaters$residuals, main='Scatterplot of residuals')
```

* In the study of tcell count verses weight transported for Black-eared wheatears.  The null hypothesis is that there is no correlation between t cell counts and weights whihc can be carried by a male black-eared wheater.

* The alternative hypothesis is that there is a correlation between t-cell counts and weight carried in black-eared wheaters.

* The assumptions required to apply a t-test on our linear regression are that the residuals are normally distributed, the relationship between the variables is linear and the datapoints are independent.  All of these assumptions are satisfied in this model.

* For a significence level of .95, the critical t-value is 2.09

* Using an t-test with 19 degrees of freedom on a linear regression model, the LMS slope is equal to 10.165, meaning for every 1% increase in t-cell ratio, the weight a wheater can carry in a season increased by 100 grams, and the intercept is equal to 3.911.  This means a wheater who has no t-cells could theoretically carry 3.911 Kg. 

* The 95% confidence interval for the slope of tcells verses carried weight is is 2.14 to 18.18.  The p-value for the t-test is .00611.

* Based on the t-test, we will reject the null hypothesis that there is no coorelation between tcell count and weight carried by Black-Eared Wheaters.  There is evidence to show a  correlation between t-cell count and weight.  Since this study was observational, we cannot assume this relationship to be causal.

## 2. Using the data in problem 1: Calculate by “hand” (using Excel) the following: (An example of this was in the PowerPoints and in the videos below.)=
### a. β ̂_(0 ) and β ̂_1
### b, The t-statistics and p-values for each.
### c. 99%  Confidence Interval Calculations for X = {3,4,5,6,7,8,9} grams.  
### d. 99% Prediction Interval Calculations for X = {3,4,5,6,7,8,9} grams.  
### e. Calibration Interval Calculations for Y ̅=0.3 and for Y = 0.3 mm.  
#### i. Graphical Method
#### ii, Analytical Method

* This has been worked out on the attached .xlsx document.

### f. Provide Plots for the Confidence Intervals and Prediction Intervals (From Excel … fully labeled!) … an example is shown for the Movie data in the videos below:

>* Beta1 and Beta0 http://screencast.com/t/ztSxTImiOk6s
>* SE of Beta1 and Beta 0 and RMSE: http://screencast.com/t/V9gnhSwb
>* Confidence Intervals: https://www.screencast.com/t/ELiUGTe7Kc
>* Prediction Intervals: https://www.screencast.com/t/ap8WETxsGUqN
>* CI and PI Plotting: https://www.screencast.com/t/efrpHrqgYZnG
>* Calibration Mean Gross:  https://www.screencast.com/t/Yu7eqiiH0X
>* Calibration Single Movie:  https://www.screencast.com/t/2vS1lGqtJ



## 3. For question 1, provide a measure of the amount of variation in the response that is accounted for by the explanatory variable.  Interpret this measure clearly indicating the units of the response and the explanatory variables. Use any method (software) you choose.

* Performing a linear regression on the Black-eared wheatear dataset shows there is a linear relationship between the tcell count of the birds, and the amount of nesting rock weight in Kilograms the birds can carry in a year.  The regression shows that for every 1% increase in t-cell response, the bird can carry an additional 100 grams of nesting rocks.  If the bird had no t-cell response, it would be able to carry 3.9 Kg in nesting eggs.  The percent of variance in nesting stone weight that is explained by the t-cell response is 33%.  This means that after applying a t-cell based linear regression model, 2/3 of the original varaince in the nesting rock weight is still present in the residuals.

 
 
 
 
 


